---
title: "A More Elaborate Example"
output: html_document
---

### Question  
Is the percentage of smokers significantly different between groups of patients?

### Data  

```{r prop_test, comment="", echo=FALSE}

smokers  <- c( 80, 84)#, 129, 90)
patients <- c( 86, 93)#, 136, 120)
smokers_percent <- paste0(round(smokers/patients * 100,1),"%")

results <- prop.test(smokers, patients)

estimates <- paste(round(results$estimate * 100,1),"%", collapse=" vs. ", sep="")

p_value <- ifelse(results$p.value < .001, 0.001, round(results$p.value,3))
p_eq_lt <- ifelse(p_value > .001,"=", "<")
p_sig <- ifelse(p_value > .001, "not", "")


```

```{r table, results='asis', echo=FALSE}
mydata <- data.frame(Smokers = smokers, 
                     Total_Patients = patients, 
                     Percent_Smokers = smokers_percent)

knitr::kable(mydata,
             row.names = TRUE,
             align = c('c', 'c', 'c'))
```

### Results
This example includes count of smokers from `r length(patients)` groups of patients. Chi-square analysis indicates the proportions of smokers are `r p_sig` significantly different between groups `r p_sig` significantly different between groups (`r estimates`, $p$-value `r paste(p_eq_lt, p_value)`).

```{r prop.test_results, echo=FALSE, comment=""}
results
```


