---
title: "Knitr chunks"
output:
  html_document:
    toc: yes
---
When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

### R code & results
```{r, include=FALSE}
summary(cars)
```

You can also embed plots, for example:

### Plots
```{r, include=FALSE, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


### Inline R code

You can insert an R expression in-line by using a single back-tick with the letter r:

```{r, results='asis', echo=FALSE}
cat("`` Example: 1 + 1 = `r 1+1` ``")
```
results in

Example: 1 + 1 = `r 1+1`

### More ellaborate example

```{r prop_test, comment=""}
smokers  <- c( 80, 84 ) #, 129 ) #, 90)
patients <- c( 86, 93 ) #, 136 ) #, 120)
results <- prop.test(smokers, patients)

p_value <- ifelse(results$p.value < .001, 0.001, round(results$p.value,3))
p_eq_lt <- ifelse(p_value > .001,"=", "<")
p_sig <- ifelse(results$p.value < .001, "", "not")

results
```

This example includes the smoking status from `r length(patients)` groups of patients. ${\chi}^2$ analysis of the proportion of smokers vs non-smokers indicate they are `r p_sig` significantly different between groups ($p$-value `r paste(p_eq_lt, p_value)`).



