---
title: "Clinical Data Analytics"
subtitle: "Type A Data Science"
author: "Rollie Parrish"
date: "2017-02-22"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      chakra: libs/remark-latest.min.js
---

# Type A & Type B Data Scientists

### Type A Data Scientist: 
> The A is for Analysis. This type is primarily 
> concerned with making sense of data or working with it in a fairly static way. 
> The Type A Data Scientist is very similar to a statistician (and may be one) 
> but knows all the practical details of working with data that aren’t taught 
> in the statistics curriculum: data cleaning, methods for dealing with very 
> large data sets, visualization, deep knowledge of a particular domain, 
> writing well about data, and so on.

### Type B Data Scientist: 
> The B is for Building. Type B Data Scientists share some statistical background 
> with Type A, but they are also very strong coders  and may be trained software 
> engineers. The Type B Data Scientist is mainly 
> interested in using data “in production.” They build models which interact 
> with users, often serving recommendations (products, people you may know, 
> ads, movies, search results).

> Michael Hochster (Quora post)


---

# "Opinionated data analysis"

Hilary Parker (StitchFix, Not So Standard Deviations podcast)

1. Reproducible Analysis
2. Version Control
3. 

---

# one

- or use the [RStudio Addin](https://rstudio.github.io/rstudioaddins/)<sup>2</sup> "Infinite Moon Reader" to live preview the slides (every time you update and save the Rmd document, the slides will be automatically reloaded in RStudio Viewer.

.footnote[
[1] 中文用户请看[这份教程](http://slides.yihui.name/xaringan/zh-CN.html)

[2] See [#2](https://github.com/yihui/xaringan/issues/2) if you do not see the template or addin in RStudio.
]

---
# remark.js vs xaringan

Some differences between using remark.js (left) and using **xaringan** (right):

.pull-left[
1. Start with a boilerplate HTML file;

1. Plain Markdown;

1. Write JavaScript to autoplay slides;

1. Manually configure MathJax;

1. Highlight code with `*`;

1. Edit Markdown source and refresh browser to see updated slides;
]

.pull-right[
1. Start with an R Markdown document;

1. R Markdown (can embed R/other code chunks);

1. Provide an option `autoplay`;

1. Highlight code with `{{}}`;

1. The RStudio addin "Infinite Moon Reader" automatically refreshes slides on changes;
]

.footnote[[*] Not really. See next page.]

---

# R Code
---
```{r comment='#'}
# a boring regression
fit = lm(dist ~ 1 + speed, data = cars)
coef(summary(fit))

```

---
# R Plots

```{r cars, fig.height=4, dev='svg'}
par(mar = c(4, 4, 1, .1))
plot(cars, pch = 19, col = 'darkgray', las = 1)
abline(fit, lwd = 2)
```

---
# Some Tips

- Slides can be automatically played if you set the `autoplay` option under `nature`, e.g. go to the next slide every 30 seconds (this can be useful for lightning talks):

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          autoplay: 30000
    ```

--

- The option `highlightLines: true` of `nature` will highlight code lines that start with `*` or are wrapped in `{{ }}`;

    ```yaml
    output:
      xaringan::moon_reader:
        nature:
          highlightLines: true
    ```

    See examples on the next page.

---

# Some Tips


.pull-left[
## Left
An example using a leading `*`:

    ```r
    if (TRUE) {
    ** message("Very important!")
    }
    ```
Output:
```r
if (TRUE) {
* message("Very important!")
}
```

This is invalid R code, so it is a plain fenced code block that is not executed.
]

.pull-right[
## Right
An example using `{{}}`:

    `r ''````{r tidy=FALSE}
    if (TRUE) {
    *{{ message("Very important!") }}
    }
    ```
Output:
```{r tidy=FALSE}
if (TRUE) {
{{ message("Very important!") }}
}
```

It is valid R code so you can run it.
]

